---
title: "covid 19"
author: "Rashonda Weathersby"
date: "3/18/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex

$H_0$:Covid 19 deaths in Korea, between Jan-May 2020 are unrelated to sex. 
$H_A$:Covid 19 there is a relationship between Jan-May 2020 to sex. 

Table of patients:
```{r}
kRdata$Sex = as.factor(kRdata$Sex)
kRdata$result = as.factor(kRdata$result)
# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

`

  





```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")
#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
Gender plays not role in covid death results. According to there is not a major differences. 

## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
According to the numerical results, men are dying more than women 2 times as much. 

## Inferential results 


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
We reject the null hypotheses due to the p-value = 0.001 and futher the odds of the male dying is 2.3 to 1. 

## Conclusions

In conclusion in Korea the probability of a male dying from covid 2.3 times greater than women is significant. 